---
title: "W271_Lab3_Final_Project"
author: "Shih Yu Chang"
date: "November 19, 2016"
output: pdf_document
---

# Effect of Financial Aid with respect to Sex for University/College Enrollment  

Higher Education Amendments of 1998 renamed SSIG as LEAP, created GEAR UP, suspended student aid eligibility for drug convictions, and added Extended Repayment. (P.L. 105-244, October 7, 1998). The changes including following: (1) Cut Stafford loan interest rates by 0.80%. Consolidation loans round up to the nearest 1/8th of a percent (previously whole percent) and capped at 8.25%. (2) Cost of attendance may now include the cost of a personal computer. (3) Excludes parents from number in college, switching it to professional judgment (PJ). (4) Adds examples of other common special circumstances that merit PJ: tuition expenses at an elementary or secondary school, medical or dental expenses not covered by insurance, unusually high child care costs, recent unemployment of a family member, the number of parents enrolled at least half-time in a degree, certificate, or other program leading to a recognized educational credential at a Title IV school, or other changes in the family's income, assets, or student's status. (5) Allows financial aid administrators the authority to refuse to certify a student's loan application on a case by case basis, so long as the school is not discriminating based on race, national origin, religion, sex, marital status, age, or disability status. This allows schools to limit the borrowing of students in specific majors or years in school. It also allows them to refuse to certify a loan if they feel that the student has no intention of repaying the loan. (6) Authorizes the establishment a loan cancellation program for teachers. (7) Authorize the US Department of Education to verify income data submitted on the FAFSA with the IRS. This measure is intended to eliminate a major source of fraud. 

We collect data from CPS Historical Time Series Tables on School Enrollment (https://www.census.gov/hhes/school/data/cps/historical/index.html). The goal is to compare 
enrollment rate for different sex of the four age-groups, i.e., 18-19, 20-21, 22-24, and 25-29, before and
after the changes of ACT in 1998.  To this end, consider the data points (from 1959
to 1998) as the first sample and the last 17 data points (from 1999 to 2015) as the second
sample. We wish to answer the following questions: 

1.  Compute the sample mean and sample covariance matrix of each sample.

2.  Are the two samples have the same covariance matrix?  

3.  Are the mean returns of the two samples equal?  
    
## Data Visualization    
    
```{r}
# Get Data 
dataDir="C:\\Data Science App\\W271\\Shih Yu\\Lab3\\Percentage_Population_Univ_Colle_Enrolled_Age_Sex.csv"
df <- read.csv(dataDir, header=TRUE)
head(df) # df[,c("A","B","E")]
#str(df)
df_m = subset(df, select=c("18_19_years_m", "X20_21_years_m", "X22_24_years_m", "X25_29_years_m"))
df_m = df_m[rev(rownames(df_m)), ]
par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
matplot(df_m, type = c("b"), pch=1,col = 1:4, xlab = "Time Index from year 1959",   ylab = "Enrollment Rate", main="Male Enrollment Rate for Different Age-Groups") 
legend("topright", inset=c(-0.3, 0), legend = expression("18 - 19", "20 - 21", "22 - 24", "25 - 29"), col=1:4, pch=c(1,3), title="Age-Groups") # optional legend

df_f = subset(df, select=c("X18_19_years_f", "X20_21_years_f", "X22_24_years_f", "X25_29_years_f"))
df_f = df_f[rev(rownames(df_f)), ]
par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
matplot(df_f, type = c("b"), pch=1,col = 1:4, xlab = "Time Index from year 1959",   ylab = "Enrollment Rate", main="Female Enrollment Rate for Different Age-Groups") 
legend("topright", inset=c(-0.3, 0), legend = expression("18 -19", "20 - 21", "22 - 24", "25 - 29"), col=1:4, pch=c(1,3), title="Age-Groups") # optional legend

```
    
## Compute the sample mean and sample covariance matrix of each sample.

```{r}
# Data for male enrollment from 1959 to 1998
x1_m = as.matrix(df[19:58, 1:4])


# Data for female enrollment from 1959 to 1998
x1_f = as.matrix(df[19:58, 5:8])

# Data for male enrollment from 1999 to 2015
x2_m = as.matrix(df[2:18, 1:4])

# Data for female enrollment from 1999 to 2015
x2_f = as.matrix(df[2:18, 5:8])

# mean of male enrollment rate from 1959 to 1998
apply(x1_m,2,mean)

# mean of female enrollment rate from 1959 to 1998
apply(x1_f,2,mean)

# mean of male enrollment rate from 1999 to 2015
apply(x2_m,2,mean)

# mean of female enrollment rate from 1999 to 2015
apply(x2_f,2,mean)

# covariance matrix of male enrollment rate from 1959 to 1998
cov(x1_m)

# covariance matrix of female enrollment rate from 1959 to 1998
cov(x1_f)

# covariance matrix of male enrollment rate from 1999 to 2015
cov(x2_m)

# covariance matrix of female enrollment rate from 1999 to 2015
cov(x2_f)

```

From mean value pespective, the male enrollment becomes lower then female at second period (from 1999 to 2015) for all age-groups. For both Sex, students enrollement rates also increase from the first period (from 1959 to 1998) to the second period (from 1999 to 2015) for all age-groups.

## Are the two samples have the same covariance matrix?  

```{r}
source("ama.R")
args(BoxM)
x_m <- rbind(x1_m,x2_m)
BoxM(x_m, c(40,17))
```

Based on the Box-M test statistic, the null hypothesis of equal covariance matrices is
rejected. The test statistic is 25.28 with p-value close to zero. For male Sex, the covariance matrix is different for the first sample period (from 1959 to 1998) and the second sample period (from 1999 to 2015).

```{r}
source("ama.R")
args(BoxM)
x_f <- rbind(x1_f,x2_f)
BoxM(x_f, c(40,17))
```

Based on the Box-M test statistic, the null hypothesis of equal covariance matrices is
rejected. The test statistic is 37.4 with p-value close to zero. For female Sex, the covariance matrix is different for the first sample period (from 1959 to 1998) and the second sample period (from 1999 to 2015).


## Are the mean enrollment of the two samples equal?

```{r}
args(cmeans)
cmeans(x_m, c(40,17),eqV=F)
```
    
Based  on  Hotelling T2 test  (with  un-equal  covariances) for male Sex,  one  can  reject  the  null
hypothesis  of  equal  mean vectors  at the  5% level.  The  test statistic  is  82.0  with
p-value 8.2e-10.    


```{r}
args(cmeans)
cmeans(x_f, c(40,17),eqV=F)
```
    
Based  on  Hotelling T2 test  (with  un-equal  covariances) for feale Sex,  one  can  reject  the  null
hypothesis  of  equal  mean vectors  at the  5% level.  The  test statistic  is  174.5  with
p-value 1.554312e-15. Comparing male and female, since femal group has larger Hotelling T2 test statistics with smaller p-value, this demonstrates that 1998 Finanicial Aid Act has Sex effect for university or college enrollment rate. 

# Time Series Analysis for future University/College Enrollment Rate  

```{r}
# df_m : data frame for male
# df_f : data frame for female
library(gplots)
library(MTS)
par(mar = rep(2, 4))
dim(df_m)
# For numerical stability
rtn = log(df_m[, 1:4] + 2)
#rtn = df_m[, 1:4]
ccm(rtn)
graphics.off()
VARorder(rtn, maxp = 5, output = T)
```

The VARorder provides informative order selection.  The cross-correlation ma-
trices, only the other hand, show that lag-1, to lag-2 have some significant cross-
correlations.  Also, the simple one-sample t-test shows that the means of the four series
are significantly different from zero.  Consequently, I start with a VAR(2) model with
constant term.  The model, however, only employs lag-1 and log-2 coefficient matrix.

```{r}
apply(rtn,2,t.test)
# Iteration 

n1=VARs(rtn, lags = c(1, 2))#, include.mean = T, fixed = diag(4), prelim = F, details = F, thres = 2)
#n1a=refVMAs(n1,thres=0.6)
```

```{r}
n1a=refVARs(n1,thres=0.5)
```

After several iterations with various choices of the threshold in model simplification, the fitted model with AIC about -24.7 is mean 

$$
\begin{bmatrix}
 0.9666  \\ 
-0.0167  \\ 
-0.7454  \\ 
1.51474  \\ 
\end{bmatrix}
$$

and AR(1) coefficients matrix as 

$$
\begin{bmatrix}
0.756 & 0.188 & -0.0718 & -0.00338 \\
0.688 & 0.391 &  0.0264 &  0.13684 \\
0.622 & 0.254 &  0.0980 &  0.30193 \\
-0.374 & 0.178 & 0.0736 &  0.43320 \\
\end{bmatrix}
$$

and AR(2) coefficients matrix as 

$$
\begin{bmatrix}
 -0.0611 & -0.0838 & 0.197 & -0.1958 \\
 -0.2300 &  0.0749 & 0.112 & -0.2490 \\
 -0.1671 & -0.0927 & 0.340 & -0.2677 \\
 -0.3769 &  0.4039 & 0.136 &  0.0572 \\
\end{bmatrix}
$$

```{r}
result = VARpred(n1, h = 10, orig = 0, Out.level = F)
```


